---
title: "Project_IV"
author: "Tian Ni"
date: "12/5/2021"
output: html_document
---

```{r}
library(recommenderlab)
library(Matrix)
library(tidyr)
library(data.table)
library(reshape2)
library(dplyr)
```

```{r}
# rating
myurl= "https://liangfgithub.github.io/MovieData/"
ratings = read.csv(paste0(myurl, 'ratings.dat?raw=true'), 
                   sep = ':',
                   colClasses = c('integer', 'NULL'), 
                   header = FALSE)
colnames(ratings) = c('UserID', 'MovieID', 'Rating', 'Timestamp')

ratings$Timestamp = NULL

# movies
movies = readLines(paste0(myurl, 'movies.dat?raw=true'))
movies = strsplit(movies, split = "::", fixed = TRUE, useBytes = TRUE)
movies = matrix(unlist(movies), ncol = 3, byrow = TRUE)
movies = data.frame(movies, stringsAsFactors = FALSE)
colnames(movies) = c('MovieID', 'Title', 'Genres')
movies$MovieID = as.integer(movies$MovieID)

# convert accented characters
movies$Title = iconv(movies$Title, "latin1", "UTF-8")

# extract year
movies$Year = as.numeric(unlist(
  lapply(movies$Title, function(x) substr(x, nchar(x)-4, nchar(x)-1))))

```


System I.
--
recommendation based on genres.

The two recommendation schemes I choose are:
1. popularity: summation of all ratings
2. popularity: The top 10 highest average rating with review number greater than 300

```{r}
# First we construct the matrix which each elements illustrate whether
# the movie(row) is this type(column) 1=yes, 0=no.
genres = as.data.frame(movies$Genres, stringsAsFactors=FALSE)
tmp = as.data.frame(tstrsplit(genres[,1], '[|]',
                              type.convert=TRUE),
                    stringsAsFactors=FALSE)
genre_list = c("Action", "Adventure", "Animation", 
               "Children's", "Comedy", "Crime",
               "Documentary", "Drama", "Fantasy",
               "Film-Noir", "Horror", "Musical", 
               "Mystery", "Romance", "Sci-Fi", 
               "Thriller", "War", "Western")
m = length(genre_list)
genre_matrix = matrix(0, nrow(movies), length(genre_list))
for(i in 1:nrow(tmp)){
  genre_matrix[i,genre_list %in% tmp[i,]]=1
}
colnames(genre_matrix) = genre_list
remove("tmp", "genres")

# 1. popularity: summation of all ratings
# Now we calculate the rate summation of each movie 
tmp1 = ratings %>% 
  group_by(MovieID) %>% 
  summarize(ratings_per_movie = n(), ave_ratings = mean(Rating), 
            sum_ratings=sum(Rating)) %>%
  inner_join(movies, by = 'MovieID')

# for each type, find the highest 10 movie and store the movie name in 
# rec matrix
rec1=matrix(NA, 180,2)
for(i in 1:18){
  index=which(genre_matrix[,i] != 0)
  sum_rating=cbind(tmp1$sum_ratings[index],index)
  index.10=sum_rating[order(sum_rating[,1],decreasing = TRUE),2][1:10]
  rec1[(10*(i-1)+1):(10*i),1:2]=cbind(movies$MovieID[index.10], movies$Title[index.10])
}

rownames(rec1)=rep(genre_list,each=10)
colnames(rec1)=c("MovieID","Title")
rec1


# 2. popularity: review number
tmp2=tmp1 %>% 
  filter(ratings_per_movie > 300) 
rec2=matrix(NA, 10,18)
for(i in 1:18){
  index=intersect(which(genre_matrix[,i] != 0),tmp2$MovieID)
  ave_rating=cbind(tmp2$ave_ratings[index],index)
  index.10=ave_rating[order(ave_rating[,1],decreasing = TRUE),2][1:10]
  rec2[,i]=movies$Title[index.10]
}
colnames(rec2)=genre_list
rec2
```


System II.
--
collaborative recommendation system

I choose UBCF and IBCF algorithm. 

```{r}
# First create the utility matrix of rating which stored as a sparse matrix
i = paste0('u', ratings$UserID)
j = paste0('m', ratings$MovieID)
x = ratings$Rating
tmp = data.frame(i, j, x, stringsAsFactors = T)
Rmat = sparseMatrix(as.integer(tmp$i), as.integer(tmp$j), x = tmp$x)
rownames(Rmat) = levels(tmp$i)
colnames(Rmat) = levels(tmp$j)
Rmat = as(Rmat,'realRatingMatrix')


e=evaluationScheme(Rmat, method="cross",  given=3, goodRating=5 )
results=evaluate(e,method = "UBCF",type="ratings",
                 parameter = list(normalize = 'Z-score', 
                                                 method = 'Cosine', 
                                                 nn = 20))
results_1=getConfusionMatrix(results)


results.2=evaluate(e,method = "IBCF",type="ratings",
                   parameter = list(normalize = 'Z-score', 
                                                 method = 'Cosine', 
                                                 k = 20))

results_2=getConfusionMatrix(results.2)
results.table=matrix(rep(NA, 20),10,2)

# Choose MSE as an evaluation metric
for(i in 1:10){
  results.table[i,1]=results_1[[i]][1,2]
  results.table[i,2]=results_2[[i]][1,2]
}

colnames(results.table)=c("UBCF","IBCF")
results.table
# From the table we could see:
# on average, the algorithm "UBCF" gives better prediction. 

```

Technical details:

Parameters in evaluate function:
normalize: the normalization method.
method: similarity matrix used in the algorithm, "Cosine" and "Pearson"
nn/k: The number of nearest neighbor I would use to calculated the target value.

Parameters in evaluation Scheme:

method:evaluation method.

given: An integer which is the number of given ratings are randomly selected for each evaluation user and presented to the recommender algorithm to calculat recommend ratings in this case.

goodRating: Rating at which an item is considered a positive for evaluation.



I use the z-score normalization to normalize the rating matrix.

The nearest neighborhood size I choose is  20.

I used Cosine similarity metric.

I still have missing values after tunning the algotithm, and I would replace this missings with the mean value of each movie over the whole train data.










